---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
This code assumes the "activity.csv" file is present on the current working directory.
```{r}
activityInIntervals = read.csv("activity.csv")
```


## What is mean total number of steps taken per day?
Using the [aggregate](http://stat.ethz.ch/R-manual/R-patched/library/stats/html/aggregate.html) function to group by day. 

```{r}
totalActivityPerDay = aggregate(activityInIntervals$steps,
                           by = list(activityInIntervals$date),
                           FUN = sum ) #Grouping by day, summing values.
mean(as.numeric(unlist(totalActivityPerDay[2]))
     ,na.rm = TRUE) #Generating the mean, ignoring NAs.
```

## What is the average daily activity pattern?
The code which generates the time series plot follows:
```{r}
dailyActivity = aggregate(as.numeric(activityInIntervals$steps),
                          by = list(activityInIntervals$interval),
                           FUN = function(x){mean(x,na.rm = TRUE)} ) 

plot(as.numeric(unlist(dailyActivity[1])),
     as.numeric(unlist(dailyActivity[2])),
     type = "l", 
     xlab = "Interval",
     ylab = "Average number of steps taken",
     main = "Average Daily Activity")     
     
```

The following code will print out the "most walked" interval
```{r}
dailyActivity[which.max(as.numeric(unlist(dailyActivity[2]))),1]
```

## Imputing missing values
1.This will count number of rows with NAs in them.
```{r}
length(which(is.na(activityInIntervals$steps)))
```
(2,3).I will replace all na's with thier 5 minutes interval mean. I chose this strategy as it will have the least effect on both 5 minute intervals mean(which will not change at all) and whole day mean.
```{r}
modifiedActivityInIntervals = activityInIntervals #first create a copy

for (i in which(is.na(activityInIntervals$steps))){
  modifiedActivityInIntervals$steps[i] <-
    dailyActivity[which(dailyActivity[1] == activityInIntervals$interval[i])[1],2] #updating missing values
}
```
4.
Comparing values before and after the change in 3.
Histogram of total number of steps a day before:
```{r}
hist(unlist(totalActivityPerDay[2]), main = "Total steps per day prior modifing")
```
After the modification:
```{r}
totalActivityPerDay = aggregate(modifiedActivityInIntervals$steps,
                           by = list(modifiedActivityInIntervals$date),
                           FUN = sum ) #updating 
hist(unlist(totalActivityPerDay[2]), main = "Total steps per day after modifing")
```
Now caclulating the new mean and median:
```{r}
mean(unlist(totalActivityPerDay[2]))
median(unlist(totalActivityPerDay[2]))
```
As you can see, all of the above has not changed at all when imputing missing values with my selected strategy.
## Are there differences in activity patterns between weekdays and weekends?
This code will add the factor variable, day, specifing is the day weekday or weekend.
```{r}

isWeekday <- function(dates){
  days = weekdays(as.Date(dates))
  days = (days == "Saturday" | days == "Sunday")
  days
  }#helper function

activityInIntervals$day <- 
  factor(isWeekday(activityInIntervals$date),labels=c("weekday","weekend")) #adding the factor variable
```
Now comparing weekdays and weekends:
```{r}
weekdayActivityInIntervals = activityInIntervals[which(activityInIntervals$day == "weekday"),]
weekdayActivity = aggregate(as.numeric(weekdayActivityInIntervals$steps),
                          by = list(weekdayActivityInIntervals$interval),
                           FUN = function(x){mean(x,na.rm = TRUE)} ) 

weekendActivityInIntervals = activityInIntervals[which(activityInIntervals$day == "weekend"),]
weekendActivity = aggregate(as.numeric(weekendActivityInIntervals$steps),
                          by = list(weekendActivityInIntervals$interval),
                           FUN = function(x){mean(x,na.rm = TRUE)} ) 
par(mfrow =c(1,2))
plot(as.numeric(unlist(weekdayActivity[1])),
     as.numeric(unlist(weekdayActivity[2])),
     type = "l", 
     xlab = "Interval",
     ylab = "Average number of steps taken",
     main = "Average Weekday Activity")   

 plot(as.numeric(unlist(weekendActivity[1])),
     as.numeric(unlist(weekendActivity[2])),
     type = "l", 
     xlab = "Interval",
     ylab = "Average number of steps taken",
     main = "Average Weekend Activity")     

     
```
